---
title: "Ch05-dates-factors"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Lubridate

Some basic time parsing. First create a string of dates and then read them as dates.

```{r}
 dt <- c("06-01-2020", "07-01-2020", "08-01-2020")
library(lubridate)

dt.n <- mdy(dt)

day(dt.n)

wday(dt.n)

#Get week days
wday(dt.n, label = T)

month(dt.n, label = T)

weekdays(dt.n)

# Doing some manipulation of the periods

dt.n + period("3 weeks")

dt.n + period("6 months")

ymd("2014-12-31")
ymd("2014-12-31") + period("7 months")

```

There can be some confusion on adding dates.
The length of months and years change so often that doing arithmetic with them can be unintuitive. Consider a simple operation, January 31st + one month. Should the answer be

February 31st (which doesn't exist)
March 4th (31 days after January 31), or
February 28th (assuming its not a leap year)
A basic property of arithmetic is that a + b - b = a. Only solution 1 obeys this property, but it is an invalid date. I've tried to make lubridate as consistent as possible by invoking the following rule if adding or subtracting a month or a year creates an invalid date, lubridate will return an NA. This is new with version 1.3.0, so if you're an old hand with lubridate be sure to remember this!

If you thought solution 2 or 3 was more useful, no problem. You can still get those results with clever arithmetic, or by using the special %m+% and %m-% operators. %m+% and %m-% automatically roll dates back to the last day of the month, should that be necessary.

```{r}
ymd("2014-12-31") + period("6 months")

ymd("2014-12-31") %m+% period("6 months")



```

This one works since adding six months to the above date would have resulted in
2015-06-31, which is invalid. `%m+%` will rollback to the last day of the month if
needed.

See this example

```{r}
jan31 <- ymd("2020-01-31")
jan31 + months(0:11)
jan31 %m+% months(0:11)
```


